---
title: 'Packageless coding: Programming with `base` `r`'
author: "Renee LaManna"
date: "11/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Objects and classes

## Vectors

### Creating a vector 

```{r}
# Numeric vector
my_vector <- vector(mode = "numeric", length = 100)
# Logical vector
my_second_vector <- vector(mode = "logical", length = 20 )
```

### Vector properties

#### Length

```{r}
# Length of my vector 
length(my_vector)
length(my_second_vector)
```
#### Class

```{r}
# What's in my vector
class(my_vector)
class(my_second_vector)
```

#### Indexing (accessing) vectors

```{r}
# Using c to create a vector, override my_vector
my_vector <-c(1:50)

# What's the first element?
my_vector[1]

# What's the last element?
my_vector[-1]

# What's the seventh element?
my_vector[7]

# Select a range from a vector
my_vector[3:7]
```

## Matrices

### Creating matrices

```{r}
# Create a 10 x 10 matrix full of NA's
my_matrix <- matrix(data = 1:100, nrow = 20, ncol = 5)
```

### Matrix properties

#### Dimensions

```{r}
# What are the dimensions
dim(my_matrix)
```

#### Length

```{r}
# How many elements are there
length(my_matrix)
```

#### Class

```{r}
class(my_matrix)
```

#### Indexing 

```{r}
# first number indicates row, if include no other number is indicated, going to assume you want all information from 2nd row, even if you dont know how big
my_matrix[2,]

# index a column
my_matrix[,3]

# Index a given cell
my_matrix[8,3]

# Index and replace
my_matrix[8,3] <- 100000

# Replace an entire column
my_matrix[,5] <- 10
```

# Control Flow

## `for` loops

1. variable ( aka counter),  2. sequence (going to repeat things/expression), 3. expression 

```{r}
for (i in 1:10){
  print("Hello world!")}
```

# Generating random numbers

```{r}
rnorm(n = 5, mean = 10, sd = 2)
# will get different values everytime you run this because it is random
```

# Let's combine all of the above

We will use `numeric` to create an empty vector
We will use `for` to iterate along items of that empty vector
We will use `[]` to access each slot and fill-in that empty vector
We will use `rnorm` to sample values from a random normal distribution 
We will plot the trend

$$N_{t-1} + r\left(1 - \frac{N_{t-1}}{K}\right)$$

$$
r \sim norm(\mu = 1.5, \sigma = 0.8)
$$

## Writing the code down

```{r}
# 1. Set up model parameters:

# Set up value of carrying capacity
K <- 100
# Define number of iterations 
nsteps <- 100
# Define mu, this is known and sd
mu <- 0.4
sigma <- 1


# 2. Set up model objects:

# Create a vector that will contain 1 element for population size of each time stamp, year 1 is first posistion of vector, year 2 in second posistion and so on
N <- vector(mode = "numeric", length = nsteps)
# Modify first element of population because right now there are not individuals (N = 0) for everything
N [1] <- 25
# Formula will then do for rest of years

# 3. Model it:

# Why start at 2? manually modified/indexed first posistion 
for(t in 2:nsteps){
  r <- rnorm(n = 1, mean = mu, sd = sigma)
  N[t] <- N[t-1] + r * (1- (N[t-1] / K))
}

# 4. Visualize it:
plot(N)
```

# Something more
We will use `matrix` to create an empty matrix (rows are time, columns are runs)
We will repeat the process